---
title: "Stat 530 Project Step 3"
author: "Yiquan Xiao"
date: "2024-04-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r import libraries, echo=FALSE}
# import libraries
library(vcd)
library(vcdExtra)
library(DescTools)
library(MASS)
library(gpairs)
library(ggplot2)
library(effects)
library(lmtest)
library(splines)
library(car)
library(carData)

library(gridGraphics)
library(gridExtra)
```


# Create Dataset
```{r prepare data}
# data
food_choice_data <- c(
  7, 1, 0, 0, 5,
  4, 0, 0, 1, 2,
  16, 3, 2, 2, 3,
  3, 0, 1, 2, 3,
  2, 2, 0, 0, 1,
  13, 7, 6, 0, 0,
  3, 9, 1, 0, 2,
  0, 1, 0, 1, 0,
  3, 7, 1, 0, 1,
  8, 6, 6, 3, 5,
  2, 4, 1, 1, 4,
  0, 1, 0, 0, 0,
  13, 10, 0, 2, 2,
  9, 0, 0, 1, 2,
  3, 9, 1, 0, 1,
  8, 1, 0, 0, 1
)
# variables and corresponding levels
Lake <- c("Hancock", "Oklawaha", "Trafford", "George")
Gender <- c("Male", "Female")
Size <- c("<= 2.3", "> 2.3")
Primary_Food_Choice <- c("Fish", "Invertebrate", "Reptile", "Bird", "Other")

# make it 4d array
food_choice_array <- array(food_choice_data, dim = c(4, 2, 2, 5),
                           dimnames = list(Lake = Lake,
                                           Gender = Gender,
                                           Size = Size,
                                           Primary_Food_Choice = Primary_Food_Choice))

# make 4-way table
food_choice_table <- as.table(food_choice_array)
ftable(food_choice_table)
```


# Log-linear Models
```{r}
food_choice_df <- as.data.frame(food_choice_table)

# baseline model: [Food][Lake Gender Size]
glm_base <- glm(Freq ~ Primary_Food_Choice + Lake * Gender * Size, 
                data = food_choice_df, family = "poisson")
LRstats(glm_base)

mosaic(glm_base, shade = TRUE, formula = ~ Primary_Food_Choice + Lake + Gender + Size, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4), 
       main = "Model: [Food][Lake Gender Size]", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)

mosaic(glm_base, shade = TRUE, formula = ~  Primary_Food_Choice + Size + Lake + Gender, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4), 
       main = "Model: [Food][Lake Gender Size]", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)
```

LRstats shows that current model is not acceptable. 
From mosaic display, we can see primary food choices of alligators vary among lakes --> association between Food and Lake


```{r}
# [Food Lake][Lake Gender Size]
glm_next1 <- glm(Freq ~ Primary_Food_Choice * Lake + Lake * Gender * Size, 
                 data = food_choice_df, family = "poisson")

LRstats(glm_next1)

cat("--------------------------------------------------------------------\n")

anova(glm_base, glm_next1, test = "Chisq")

mosaic(glm_next1, shade = TRUE, formula = ~ Primary_Food_Choice + Lake + Gender + Size, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4), 
       main = "Model: [Food Lake][Lake Gender Size]", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)
```

anova confirms that there is an association between Food and Lake. 
LRstats shows that current model is not acceptable. 


```{r}
# [Food Lake][Food Size][Lake Gender Size]
glm_next2.1 <- glm(Freq ~ Primary_Food_Choice * (Lake + Size) + Lake * Gender * Size, 
                   data = food_choice_df, family = "poisson")

LRstats(glm_next2.1)

cat("--------------------------------------------------------------------\n")

anova(glm_next1, glm_next2.1, test = "Chisq")

mosaic(glm_next2.1, shade = TRUE, formula = ~ Primary_Food_Choice + Lake + Gender + Size, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4), 
       main = "Model: [Food Lake][Food Size][Lake Gender Size]", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)
```


```{r}
# [Food Lake Size][Lake Gender Size]
glm_next2.2 <- glm(Freq ~ Primary_Food_Choice * Lake * Size + Lake * Gender * Size, 
                   data = food_choice_df, family = "poisson")

LRstats(glm_next2.2)

cat("--------------------------------------------------------------------\n")

anova(glm_next2.1, glm_next2.2, test = "Chisq")

mosaic(glm_next2.2, shade = TRUE, formula = ~ Primary_Food_Choice + Lake + Gender + Size, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4), 
       main = "Model: [Food Lake Size][Lake Gender Size]", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)

mosaic(glm_next2.2, shade = TRUE, formula = ~ Lake + Primary_Food_Choice + Size + Gender, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4), 
       main = "Model: [Food Lake Size][Lake Gender Size]", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)
```






```{r}
# [Food Lake Size][Food Gender][Lake Gender Size]
glm_next3.1 <- glm(Freq ~ Primary_Food_Choice * (Lake * Size + Gender) + Lake * Gender * Size, 
                   data = food_choice_df, family = "poisson")

LRstats(glm_next3.1)

cat("--------------------------------------------------------------------\n")

anova(glm_next2.2, glm_next3.1, test = "Chisq")

mosaic(glm_next3.1, shade = TRUE, formula = ~ Primary_Food_Choice + Lake + Gender + Size, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4), 
       main = "Model: [Food Lake Size][Food Gender][Lake Gender Size]", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)
```


```{r}
# [Food Lake Size][Food Lake Gender][Lake Gender Size]
glm_next3.2 <- glm(Freq ~ Primary_Food_Choice * (Lake * Size + Lake * Gender) 
                   + Lake * Gender * Size, 
                   data = food_choice_df, family = "poisson")

LRstats(glm_next3.2)

cat("--------------------------------------------------------------------\n")

anova(glm_next3.1, glm_next3.2, test = "Chisq")

mosaic(glm_next3.2, shade = TRUE, formula = ~ Primary_Food_Choice + Lake + Gender + Size, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4), 
       main = "Model: [Food Lake Size][Food Lake Gender][Lake Gender Size]", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)
```



```{r}
# [Food Lake Size][Food Gender Size][Lake Gender Size]
glm_next3.3 <- glm(Freq ~ Primary_Food_Choice * (Lake * Size + Gender * Size) 
                   + Lake * Gender * Size, 
                   data = food_choice_df, family = "poisson")

LRstats(glm_next3.3)

cat("--------------------------------------------------------------------\n")

anova(glm_next3.3, glm_next3.2, test = "Chisq")

cat("--------------------------------------------------------------------\n")

loglm_next3.2 <- loglm(~ ~ Primary_Food_Choice * (Lake * Size + Lake * Gender) 
                       + Lake * Gender * Size, 
                       data = food_choice_table)

loglm_next3.3 <- loglm(~ Primary_Food_Choice * (Lake * Size + Gender * Size) 
                       + Lake * Gender * Size, 
                       data = food_choice_table)

anova(loglm_next3.3, loglm_next3.2, test = "Chisq")

mosaic(glm_next3.3, shade = TRUE, formula = ~ Primary_Food_Choice + Lake + Gender + Size, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4), 
       main = "Model: [Food Lake Size][Food Gender Size][Lake Gender Size]", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)
```


```{r}
# [Food Lake Gender Size]
glm_sat <- glm(Freq ~ Primary_Food_Choice * Lake * Gender * Size, 
                   data = food_choice_df, family = "poisson")

LRstats(glm_sat)

cat("--------------------------------------------------------------------\n")

anova(glm_next3.3, glm_sat, test = "Chisq")

mosaic(glm_sat, shade = TRUE, formula = ~ Primary_Food_Choice + Lake + Gender + Size, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4), 
       main = "Model: [Food Lake Gender Size]", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)

```


```{r}
# [Food Lake Size][Food Gender Size]
glm_next4.1 <- glm(Freq ~ Primary_Food_Choice * (Lake * Size + Gender * Size), 
                   data = food_choice_df, family = "poisson")

LRstats(glm_next4.1)

cat("--------------------------------------------------------------------\n")

anova(glm_next4.1, glm_next3.3, test = "Chisq")

mosaic(glm_next4.1, shade = TRUE, formula = ~ Primary_Food_Choice + Lake + Gender + Size, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4), 
       main = "Model: [Food Lake Size][Food Gender Size]", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)
```


```{r}
# [Food Lake Size][Food Gender Size][Lake Gender]
glm_next4.2 <- glm(Freq ~ Primary_Food_Choice * (Lake * Size + Gender * Size) + Lake * Gender, 
                   data = food_choice_df, family = "poisson")

LRstats(glm_next4.2)

cat("--------------------------------------------------------------------\n")

anova(glm_next4.2, glm_next3.3, test = "Chisq")

mosaic(glm_next4.2, shade = TRUE, formula = ~ Primary_Food_Choice + Lake + Gender + Size, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4),
       main = "Model: [Food Lake Size][Food Gender Size][Lake Gender]", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)
```


```{r}
food_choice_df <- within(food_choice_df,
                         {
                           FO_LH_GF_SG <-
                             (Primary_Food_Choice == "Other") *
                             (Lake == "Hancock") *
                             (Gender == "Female") *
                             (Size == "> 2.3")
                         }
)


glm_next4.2.1 <- glm(Freq ~ Primary_Food_Choice * (Lake * Size + Gender * Size) 
                     + Lake * Gender 
                     + FO_LH_GF_SG, 
                     data = food_choice_df, family = "poisson")

LRstats(glm_next4.2.1)

cat("--------------------------------------------------------------------\n")

anova(glm_next4.2, glm_next4.2.1, test = "Chisq")

mosaic(glm_next4.2.1, shade = TRUE, formula = ~ Primary_Food_Choice + Lake + Gender + Size, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4),
       # main = "Model: [Food Lake Size][Food Gender Size][Lake Gender]", 
       main = "Model: [FLS][FGS][LG] with 1 extra term", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)

```


```{r}
food_choice_df <- within(food_choice_df,
                         {
                           FI_LH_GF_SL <-
                             (Primary_Food_Choice == "Invertebrate") *
                             (Lake == "Hancock") *
                             (Gender == "Female") *
                             (Size == "<= 2.3")
                         }
)

food_choice_df <- within(food_choice_df,
                         {
                           FF_LH_GF_SG <-
                             (Primary_Food_Choice == "Fish") *
                             (Lake == "Hancock") *
                             (Gender == "Female") *
                             (Size == "> 2.3")
                         }
)


glm_next4.2.2 <- glm(Freq ~ Primary_Food_Choice * (Lake * Size + Gender * Size) 
                     + Lake * Gender 
                     + FO_LH_GF_SG 
                     + FI_LH_GF_SL 
                     + FF_LH_GF_SG, 
                     data = food_choice_df, family = "poisson")

LRstats(glm_next4.2.2)

cat("--------------------------------------------------------------------\n")

anova(glm_next4.2.1, glm_next4.2.2, test = "Chisq")

mosaic(glm_next4.2.2, shade = TRUE, formula = ~ Primary_Food_Choice + Lake + Gender + Size, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4),
       main = "Model: [FLS][FGS][LG] with 3 extra term", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)

```


```{r}
food_choice_df <- within(food_choice_df,
                         {
                           FI_LO_GF_SL <-
                             (Primary_Food_Choice == "Invertebrate") *
                             (Lake == "Oklawaha") *
                             (Gender == "Female") *
                             (Size == "<= 2.3")
                         }
)

food_choice_df <- within(food_choice_df,
                         {
                           FF_LO_GF_SG <-
                             (Primary_Food_Choice == "Fish") *
                             (Lake == "Oklawaha") *
                             (Gender == "Female") *
                             (Size == "> 2.3")
                         }
)


glm_next4.2.3 <- glm(Freq ~ Primary_Food_Choice * (Lake * Size + Gender * Size) 
                     + Lake * Gender 
                     + FO_LH_GF_SG 
                     + FI_LH_GF_SL 
                     + FF_LH_GF_SG
                     + FI_LO_GF_SL 
                     + FF_LO_GF_SG, 
                     data = food_choice_df, family = "poisson")

LRstats(glm_next4.2.3)

cat("--------------------------------------------------------------------\n")

anova(glm_next4.2.2, glm_next4.2.3, test = "Chisq")

cat("--------------------------------------------------------------------\n")

anova(glm_next4.2.3, glm_sat, test = "Chisq")

mosaic(glm_next4.2.3, shade = TRUE, formula = ~ Primary_Food_Choice + Lake + Gender + Size, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4),
       main = "Model: [FLS][FGS][LG] with 5 extra term (Final Model)", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)

```


```{r}
food_choice_df <- within(food_choice_df,
                         {
                           FO_LO_GM_SL <-
                             (Primary_Food_Choice == "Other") *
                             (Lake == "Oklawaha") *
                             (Gender == "Male") *
                             (Size == "<= 2.3")
                         }
)
food_choice_df <- within(food_choice_df,
                         {
                           FI_LG_GF_SG <-
                             (Primary_Food_Choice == "Invertebrate") *
                             (Lake == "George") *
                             (Gender == "Female") *
                             (Size == "> 2.3")
                         }
)

food_choice_df <- within(food_choice_df,
                         {
                           FI_LO_GF_SG <-
                             (Primary_Food_Choice == "Invertebrate") *
                             (Lake == "Oklawaha") *
                             (Gender == "Female") *
                             (Size == "> 2.3")
                         }
)

glm_next4.1.1 <- glm(Freq ~ Primary_Food_Choice * (Lake * Size + Gender * Size) 
                     + FO_LO_GM_SL 
                     + FI_LG_GF_SG 
                     + FI_LO_GF_SG, 
                     data = food_choice_df, family = "poisson")

LRstats(glm_next4.1.1)

cat("--------------------------------------------------------------------\n")

anova(glm_next4.1, glm_next4.1.1, test = "Chisq")

mosaic(glm_next4.1.1, shade = TRUE, formula = ~ Primary_Food_Choice + Lake + Gender + Size, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4), 
       main = "Model: [Food Lake Size][Food Gender Size]", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)

```


```{r}
food_choice_df <- within(food_choice_df,
                          {
                            FO_LH_GM_SL <-
                              (Primary_Food_Choice == "Other") *
                              (Lake == "Hancock") *
                              (Gender == "Male") *
                              (Size == "<= 2.3")
                          }
)

glm_next4.1.2 <- glm(Freq ~ Primary_Food_Choice * (Lake * Size + Gender * Size) 
                     + FO_LO_GM_SL 
                     + FI_LG_GF_SG 
                     + FI_LO_GF_SG 
                     + FO_LH_GM_SL, 
                     data = food_choice_df, family = "poisson")

LRstats(glm_next4.1.2)

cat("--------------------------------------------------------------------\n")

anova(glm_next4.1.1, glm_next4.1.2, test = "Chisq")

cat("--------------------------------------------------------------------\n")

anova(glm_next4.1.2, glm_next4.2.3, test = "Chisq")

mosaic(glm_next4.1.2, shade = TRUE, formula = ~ Primary_Food_Choice + Lake + Gender + Size, 
       split = TRUE,
       # residuals_type = "rstandard", 
       # gp = shading_Friendly, 
       # gp_args = list(interpolate = 1 : 4), 
       main = "Model: [Food Lake Size][Food Gender Size]", 
       abbreviate_labs = c(Primary_Food_Choice = TRUE, 
                           Lake = TRUE, Gender = TRUE, Size = TRUE),
       keep_aspect_ratio = FALSE)

```



# Logistic Regression

## Response: Fish
```{r}

fish_data <- food_choice_array[, , , 1]
not_fish_data <- apply(food_choice_array[, , , -1], c(1, 2, 3), sum)

new_food_choice_data <- array(c(not_fish_data, fish_data), dim = c(4, 2, 2, 2))
dimnames(new_food_choice_data) <- list(
  Lake = c("Hancock", "Oklawaha", "Trafford", "George"),
  Gender = c("Male", "Female"),
  Size = c("<= 2.3", "> 2.3"),
  Fish = c("Not Fish", "Fish")
)

fish_table <- as.table(new_food_choice_data)
ftable(fish_table)

```


```{r turn to case dfm}

# Convert table to dataframe
fish_dfm <- as.data.frame(fish_table)

# Names for expanding grid based on the table's dimnames
names_vector <- dimnames(fish_table)

# Create all possible combinations of Lake, Gender, Size, and Fish
combination_df <- expand.grid(names_vector)

# Include frequencies from the table
combination_df$Freq <- c(fish_table)

# Expand the dataframe based on the 'Freq' column
fish_dfm <- combination_df[rep(seq_len(nrow(combination_df)), combination_df$Freq), 1:4]  # Exclude the Freq column

# Reset row names to avoid duplicate row names
row.names(fish_dfm) <- NULL

# View the fish_dfm dataframe
fish_dfm

```



```{r}

fish.mod0 <- glm(Fish ~ Lake + Gender + Size,
                 data = fish_dfm, family =binomial)
Anova(fish.mod0)

cat("-----------------------------------------------------------------------\n")

fish.mod1 <- glm(Fish ~ (Lake + Gender + Size)^2,
                 data = fish_dfm, family =binomial)
Anova(fish.mod1)

cat("-----------------------------------------------------------------------\n")

anova(fish.mod0, fish.mod1, test = "Chisq")

cat("=======================================================================\n")

fish.mod2 <- glm(Fish ~ Lake * Gender * Size,
                 data = fish_dfm, family =binomial)
Anova(fish.mod2)

cat("-----------------------------------------------------------------------\n")

anova(fish.mod1, fish.mod2, test = "Chisq")

cat("=======================================================================\n")

LRstats(fish.mod1)
```


```{r}

fish.mod3 <- glm(Fish ~ Lake * Size + Gender,
                 data = fish_dfm, family =binomial)
Anova(fish.mod3)

cat("=======================================================================\n")

LRstats(fish.mod3)

cat("=======================================================================\n")

anova(fish.mod3, fish.mod1, test = "Chisq")
```


```{r}
coeftest(fish.mod3)
```


```{r}
fish.eff3 <- allEffects(fish.mod3)

plot(fish.eff3, rows = 1, cols = 2, type="response")
```


```{r}
plot_list <- list()

plot(Effect("Gender", fish.mod3), lwd = 3, multiline = TRUE, ci.style = "bands",
     type="response",  xlab = "Gender", ylab = "Fish", 
     key.args = list(x = .05, y = .99, cex = 1.2, columns = 1))
plot_list[[1]] <- grid.grab(wrap.grobs = TRUE)

plot(Effect(c("Lake", "Size"), fish.mod3), lwd = 3, multiline = TRUE, ci.style = "bands", 
     type="response",  xlab = "Lake", ylab = "Fish", ylim = c(-0.05, 0.8), 
     key.args = list(x = .05, y = .99, cex = 1.2, columns = 1))
plot_list[[2]] <- grid.grab(wrap.grobs = TRUE)

grid.arrange(grobs = plot_list, ncol = 2, nrow = 1)
```



## Response: Invertabrate
```{r}

invertebrate_data <- food_choice_array[, , , 2]
not_invertebrate_data <- apply(food_choice_array[, , , -2], c(1, 2, 3), sum)

new_food_choice_data <- array(c(not_invertebrate_data, invertebrate_data), dim = c(4, 2, 2, 2))
dimnames(new_food_choice_data) <- list(
  Lake = c("Hancock", "Oklawaha", "Trafford", "George"),
  Gender = c("Male", "Female"),
  Size = c("<= 2.3", "> 2.3"),
  Invertebrate = c("Not Invertebrate", "Invertebrate")
)

invertebrate_table <- as.table(new_food_choice_data)
ftable(invertebrate_table)

```


```{r turn to case dfm}

# Convert table to dataframe
invertebrate_dfm <- as.data.frame(invertebrate_table)

# Names for expanding grid based on the table's dimnames
names_vector <- dimnames(invertebrate_table)

# Create all possible combinations of Lake, Gender, Size, and Invertebrate
combination_df <- expand.grid(names_vector)

# Include frequencies from the table
combination_df$Freq <- c(invertebrate_table)

# Expand the dataframe based on the 'Freq' column
invertebrate_dfm <- combination_df[rep(seq_len(nrow(combination_df)), combination_df$Freq), 1:4]  # Exclude the Freq column

# Reset row names to avoid duplicate row names
row.names(invertebrate_dfm) <- NULL

# View the invertebrate_dfm dataframe
invertebrate_dfm

```


```{r}

invertebrate.mod0 <- glm(Invertebrate ~ Lake + Gender + Size,
                         data = invertebrate_dfm, family = "binomial")
Anova(invertebrate.mod0)

cat("-----------------------------------------------------------------------\n")

invertebrate.mod1 <- glm(Invertebrate ~ (Lake + Gender + Size)^2,
                         data = invertebrate_dfm, family = "binomial")
Anova(invertebrate.mod1)

cat("-----------------------------------------------------------------------\n")

anova(invertebrate.mod0, invertebrate.mod1, test = "Chisq")

cat("=======================================================================\n")

invertebrate.mod2 <- glm(Invertebrate ~ Lake * Gender * Size,
                         data = invertebrate_dfm, family = "binomial")
Anova(invertebrate.mod2)

cat("-----------------------------------------------------------------------\n")

anova(invertebrate.mod0, invertebrate.mod2, test = "Chisq")

cat("=======================================================================\n")

LRstats(invertebrate.mod2)
```


```{r}
coeftest(invertebrate.mod2)
```


```{r}
invertebrate.eff2 <- allEffects(invertebrate.mod2)

plot(invertebrate.eff2, rows = 1, cols = 1, type="response")
```


```{r}

plot(Effect(c("Lake", "Gender", "Size"), invertebrate.mod2), lwd = 3, 
     multiline = TRUE, ci.style = "bands", 
     type="response",  xlab = "Lake", ylab = "Invertebrate", ylim = c(-0.05, 0.7), 
     key.args = list(x = .05, y = .95, cex = 1.2, columns = 1))

```


## Response: Reptile
```{r}

reptile_data <- food_choice_array[, , , 3]
not_reptile_data <- apply(food_choice_array[, , , -3], c(1, 2, 3), sum)

new_food_choice_data <- array(c(not_reptile_data, reptile_data), dim = c(4, 2, 2, 2))
dimnames(new_food_choice_data) <- list(
  Lake = c("Hancock", "Oklawaha", "Trafford", "George"),
  Gender = c("Male", "Female"),
  Size = c("<= 2.3", "> 2.3"),
  Reptile = c("Not Reptile", "Reptile")
)

reptile_table <- as.table(new_food_choice_data)
ftable(reptile_table)

```


```{r turn to case dfm}

# Convert table to dataframe
reptile_dfm <- as.data.frame(reptile_table)

# Names for expanding grid based on the table's dimnames
names_vector <- dimnames(reptile_table)

# Create all possible combinations of Lake, Gender, Size, and Reptile
combination_df <- expand.grid(names_vector)

# Include frequencies from the table
combination_df$Freq <- c(reptile_table)

# Expand the dataframe based on the 'Freq' column
reptile_dfm <- combination_df[rep(seq_len(nrow(combination_df)), combination_df$Freq), 1:4]  # Exclude the Freq column

# Reset row names to avoid duplicate row names
row.names(reptile_dfm) <- NULL

# View the reptile_dfm dataframe
reptile_dfm

```


```{r}

reptile.mod0 <- glm(Reptile ~ Lake + Gender + Size,
                    data = reptile_dfm, family = "binomial")
Anova(reptile.mod0)

cat("-----------------------------------------------------------------------\n")

reptile.mod1 <- glm(Reptile ~ (Lake + Gender + Size)^2,
                    data = reptile_dfm, family = "binomial")
Anova(reptile.mod1)

cat("-----------------------------------------------------------------------\n")

anova(reptile.mod0, reptile.mod1, test = "Chisq")

cat("=======================================================================\n")

reptile.mod2 <- glm(Reptile ~ Lake * Gender * Size,
                    data = reptile_dfm, family = "binomial")
Anova(reptile.mod2)

cat("-----------------------------------------------------------------------\n")

anova(reptile.mod1, reptile.mod2, test = "Chisq")

cat("=======================================================================\n")

LRstats(reptile.mod1)

```


```{r}
reptile.mod3 <- glm(Reptile ~ Lake * Gender + Lake * Size,
                    data = reptile_dfm, family =binomial)
Anova(reptile.mod3)

cat("=======================================================================\n")

LRstats(reptile.mod3)

cat("=======================================================================\n")

anova(reptile.mod3, reptile.mod1, test = "Chisq")
```


```{r}
coeftest(reptile.mod3)
```


```{r}
reptile.eff3 <- allEffects(reptile.mod3)

plot(reptile.eff3, rows = 1, cols = 2, type="response")
```


```{r}
plot_list <- list()

plot(Effect(c("Gender", "Lake"), reptile.mod3), lwd = 3, multiline = TRUE, ci.style = "bands", 
     type="response",  xlab = "Lake", ylab = "Reptile", ylim = c(-0.05, 0.5), 
     key.args = list(x = .05, y = .99, cex = 1.2, columns = 1))
plot_list[[1]] <- grid.grab(wrap.grobs = TRUE)

plot(Effect(c("Size", "Lake"), reptile.mod3), lwd = 3, multiline = TRUE, ci.style = "bands", 
     type="response",  xlab = "Lake", ylab = "Reptile", ylim = c(-0.05, 0.5), 
     key.args = list(x = .05, y = .99, cex = 1.2, columns = 1))
plot_list[[2]] <- grid.grab(wrap.grobs = TRUE)

grid.arrange(grobs = plot_list, ncol = 2, nrow = 1)

```


## Response: Bird
```{r}

bird_data <- food_choice_array[, , , 4]
not_bird_data <- apply(food_choice_array[, , , -4], c(1, 2, 3), sum)

new_food_choice_data <- array(c(not_bird_data, bird_data), dim = c(4, 2, 2, 2))
dimnames(new_food_choice_data) <- list(
  Lake = c("Hancock", "Oklawaha", "Trafford", "George"),
  Gender = c("Male", "Female"),
  Size = c("<= 2.3", "> 2.3"),
  Bird = c("Not Bird", "Bird")
)

bird_table <- as.table(new_food_choice_data)
ftable(bird_table)

```


```{r turn to case dfm}

# Convert table to dataframe
bird_dfm <- as.data.frame(bird_table)

# Names for expanding grid based on the table's dimnames
names_vector <- dimnames(bird_table)

# Create all possible combinations of Lake, Gender, Size, and Bird
combination_df <- expand.grid(names_vector)

# Include frequencies from the table
combination_df$Freq <- c(bird_table)

# Expand the dataframe based on the 'Freq' column
bird_dfm <- combination_df[rep(seq_len(nrow(combination_df)), combination_df$Freq), 1:4]  # Exclude the Freq column

# Reset row names to avoid duplicate row names
row.names(bird_dfm) <- NULL

# View the bird_dfm dataframe
bird_dfm

```


```{r}

bird.mod0 <- glm(Bird ~ Lake + Gender + Size,
                 data = bird_dfm, family =binomial)
Anova(bird.mod0)

cat("-----------------------------------------------------------------------\n")

bird.mod1 <- glm(Bird ~ (Lake + Gender + Size)^2,
                 data = bird_dfm, family =binomial)
Anova(bird.mod1)

cat("-----------------------------------------------------------------------\n")

anova(bird.mod0, bird.mod1, test = "Chisq")

cat("=======================================================================\n")

bird.mod2 <- glm(Bird ~ Lake * Gender * Size,
                 data = bird_dfm, family =binomial)
Anova(bird.mod2)

cat("-----------------------------------------------------------------------\n")

anova(bird.mod1, bird.mod2, test = "Chisq")

cat("=======================================================================\n")

LRstats(bird.mod1)
```


```{r}
bird.mod3 <- glm(Bird ~ Lake * Size + Gender * Size,
                 data = bird_dfm, family =binomial)
Anova(bird.mod3)

cat("=======================================================================\n")

LRstats(bird.mod3)

cat("=======================================================================\n")

anova(bird.mod3, bird.mod1, test = "Chisq")
```


```{r}
coeftest(bird.mod3)
```


```{r}
bird.eff3 <- allEffects(bird.mod3)

plot(bird.eff3, rows = 1, cols = 2, type="response")
```


```{r}
plot_list <- list()

plot(Effect(c("Size", "Lake"), bird.mod3), lwd = 3, multiline = TRUE, ci.style = "bands", 
     type="response",  xlab = "Lake", ylab = "Bird", ylim = c(-0.05, 0.6), 
     key.args = list(x = .05, y = .99, cex = 1.2, columns = 1))
plot_list[[1]] <- grid.grab(wrap.grobs = TRUE)

plot(Effect(c("Size", "Gender"), bird.mod3), lwd = 3, multiline = TRUE, ci.style = "bands", 
     type="response",  xlab = "Size", ylab = "Bird", ylim = c(-0.05, 0.6), 
     key.args = list(x = .05, y = .99, cex = 1.2, columns = 1))
plot_list[[2]] <- grid.grab(wrap.grobs = TRUE)

grid.arrange(grobs = plot_list, ncol = 2, nrow = 1)

```


## Response: Other
```{r}

other_data <- food_choice_array[, , , 5]
not_other_data <- apply(food_choice_array[, , , -5], c(1, 2, 3), sum)

new_food_choice_data <- array(c(not_other_data, other_data), dim = c(4, 2, 2, 2))
dimnames(new_food_choice_data) <- list(
  Lake = c("Hancock", "Oklawaha", "Trafford", "George"),
  Gender = c("Male", "Female"),
  Size = c("<= 2.3", "> 2.3"),
  Other = c("Not Other", "Other")
)

other_table <- as.table(new_food_choice_data)
ftable(other_table)

```


```{r turn to case dfm}

# Convert table to dataframe
other_dfm <- as.data.frame(other_table)

# Names for expanding grid based on the table's dimnames
names_vector <- dimnames(other_table)

# Create all possible combinations of Lake, Gender, Size, and Other
combination_df <- expand.grid(names_vector)

# Include frequencies from the table
combination_df$Freq <- c(other_table)

# Expand the dataframe based on the 'Freq' column
other_dfm <- combination_df[rep(seq_len(nrow(combination_df)), combination_df$Freq), 1:4]  # Exclude the Freq column

# Reset row names to avoid duplicate row names
row.names(other_dfm) <- NULL

# View the other_dfm dataframe
other_dfm

```


```{r}

other.mod0 <- glm(Other ~ Lake + Gender + Size,
                  data = other_dfm, family =binomial)
Anova(other.mod0)

cat("-----------------------------------------------------------------------\n")

other.mod1 <- glm(Other ~ (Lake + Gender + Size)^2,
                  data = other_dfm, family =binomial)
Anova(other.mod1)

cat("-----------------------------------------------------------------------\n")

anova(other.mod0, other.mod1, test = "Chisq")

cat("=======================================================================\n")

other.mod2 <- glm(Other ~ Lake * Gender * Size,
                  data = other_dfm, family =binomial)
Anova(other.mod2)

cat("-----------------------------------------------------------------------\n")

anova(other.mod1, other.mod2, test = "Chisq")

cat("=======================================================================\n")

LRstats(other.mod2)

```


```{r}
coeftest(other.mod2)
```


```{r}
other.eff2 <- allEffects(other.mod2)

plot(other.eff2, rows = 1, cols = 1, type="response")
```


```{r}

plot(Effect(c("Gender", "Size", "Lake"), other.mod2), lwd = 3, 
     multiline = TRUE, ci.style = "bands", 
     type="response",  xlab = "Lake", ylab = "Other", ylim = c(-0.05, 1.05), 
     key.args = list(x = .05, y = .95, cex = 1.2, columns = 1))

```





