---
title: "Stat 520 Project Part 2"
author: "Yiquan Xiao"
date: "2024-03-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r import libraries, echo=FALSE}
# import libraries
library(vcd)
library(vcdExtra)
library(DescTools)
```


# Create Dataset
```{r prepare data}
# data
food_choice_data <- c(
  7, 1, 0, 0, 5,
  4, 0, 0, 1, 2,
  16, 3, 2, 2, 3,
  3, 0, 1, 2, 3,
  2, 2, 0, 0, 1,
  13, 7, 6, 0, 0,
  3, 9, 1, 0, 2,
  0, 1, 0, 1, 0,
  3, 7, 1, 0, 1,
  8, 6, 6, 3, 5,
  2, 4, 1, 1, 4,
  0, 1, 0, 0, 0,
  13, 10, 0, 2, 2,
  9, 0, 0, 1, 2,
  3, 9, 1, 0, 1,
  8, 1, 0, 0, 1
)

# variables and corresponding levels
Lake <- c("Hancock", "Oklawaha", "Trafford", "George")
Gender <- c("Male", "Female")
Size <- c("<= 2.3", "> 2.3")
Primary_Food_Choice <- c("Fish", "Invertebrate", "Reptile", "Bird", "Other")

# make it 4d array
food_choice_array <- array(food_choice_data, dim = c(4, 2, 2, 5),
                           dimnames = list(Lake = Lake,
                                           Gender = Gender,
                                           Size = Size,
                                           Primary_Food_Choice = Primary_Food_Choice))

# make 4-way table
food_choice_table <- as.table(food_choice_array)
ftable(food_choice_table)
```


# Marginal Distribution of Primary Food Choice
```{r}
li_data <- c("L1", "L2", "L3", "L4", "L5", "L6", "L7", "L8", 
             "L9", "L10", "L11", "L12", "L13", "L14", "L15", "L16")

# make it 4d array
li_array <- array(li_data, dim = c(4, 2, 2),
                  dimnames = list(Lake = Lake,
                                  Gender = Gender,
                                  Size = Size))

# make 4-way table
li_table <- as.table(li_array)
ftable(li_table)
```


```{r}
generate_ltable <- function(lake_choice, gender_choice, size_choice) {
  # Initialize an empty list to store the frequency tables for each food choice
  l_frequency_tables <- list()
  
  # Loop through each primary food choice to construct the frequency table
  for (food_choice in dimnames(food_choice_table)$Primary_Food_Choice) {
    # Extract the sub-table for the current food choice
    food_choice_subtable <- food_choice_table[,,,food_choice]
    
    # Initialize a vector to store counts of L = 0, 1, 2, 3
    # l_counts <- integer(4)
    l_counts <- rep(0, 4)
    
    # Calculate L for each possible combination of Lake, Gender, and Size
    for (lake in 1:length(dimnames(food_choice_subtable)$Lake)) {
      for (gender in 1:length(dimnames(food_choice_subtable)$Gender)) {
        for (size in 1:length(dimnames(food_choice_subtable)$Size)) {
          # Calculate L based on conditions 
          l_value <- 0
          if (dimnames(food_choice_subtable)$Lake[lake] == lake_choice) l_value <- l_value + 1
          if (dimnames(food_choice_subtable)$Gender[gender] == gender_choice) l_value <- l_value + 1
          if (dimnames(food_choice_subtable)$Size[size] == size_choice) l_value <- l_value + 1
          
          # Update the count of L in the vector
          l_counts[l_value + 1] <- l_counts[l_value + 1] + 
            food_choice_subtable[lake, gender, size]
        }
      }
    }
    
    # Store the frequency table for the current food choice in the list
    l_frequency_tables[[food_choice]] <- as.table(setNames(l_counts, 0:3))
  }
  return(l_frequency_tables)
}
```


```{r}
analysis_ltable <- function(l_frequency_tables, number_of_l) {
  for (food_choice in dimnames(food_choice_table)$Primary_Food_Choice) {
    cat(food_choice, "\n")
    for (distribution in c("binomial", "poisson", "nbinomial")) {
      
      if (distribution == "binomial") {
        l.goodfit <- goodfit(l_frequency_tables[[food_choice]], 
                              type = distribution, par = list(size = 3))
      }
      else{
        l.goodfit <- goodfit(l_frequency_tables[[food_choice]], 
                              type = distribution)
      }
      
      summary(l.goodfit)
      
      plot_title = sprintf("Hanging Rootogram for fitting L%s with %s \n when Primary Food Type = %s", 
                           number_of_l, distribution, food_choice)
      
      # Hanging Rootgram
      plot(l.goodfit, type = "hanging", shade = TRUE, 
           main = plot_title, main_gp = gpar(fontsize = 10))
      
      plot_title = sprintf("Binomialness plot for fiiting L%s with %s \n when Primary Food Type = %s", 
                           number_of_l, distribution, food_choice)
      
      # Generalized Distribution plot
      if (distribution == "binomial") {
        distplot(l_frequency_tables[[food_choice]], type = distribution, size = 3,
                 main = plot_title,
                 xlab = "Number of satisfied conditions")
      }
      else{
        distplot(l_frequency_tables[[food_choice]], type = distribution,
                 main = plot_title,
                 xlab = "Number of satisfied conditions")
      }
      
    }
    cat("---------------------------------------------------------------\n")
  }
}
```


```{r}
# the degree of freedom in the test above (when Primary Food Type = Bird above) is 
# different from others because there is a "0" in the table

test_table <- as.table(setNames(c(1, 2, 3, 4), 0:3))
test_table2 <- as.table(setNames(c(1, 0, 3, 4), 0:3))

summary(goodfit(test_table, type = "nbinomial"))
summary(goodfit(test_table2, type = "nbinomial"))
```


```{r L1}
l1_frequency_tables <- generate_ltable("Hancock", "Male", "<= 2.3")
l1_frequency_tables

cat("===================================================================\n")

analysis_ltable(l1_frequency_tables, 1)
```


```{r L2}
l2_frequency_tables <- generate_ltable("Oklawaha", "Male", "<= 2.3")
l2_frequency_tables

cat("===================================================================\n")

analysis_ltable(l2_frequency_tables, 2)
```


```{r L3}
l3_frequency_tables <- generate_ltable("Trafford", "Male", "<= 2.3")
l3_frequency_tables

cat("===================================================================\n")

analysis_ltable(l3_frequency_tables, 3)
```


```{r L4}
l4_frequency_tables <- generate_ltable("George", "Male", "<= 2.3")
l4_frequency_tables

cat("===================================================================\n")

analysis_ltable(l4_frequency_tables, 4)
```


```{r L5}
l5_frequency_tables <- generate_ltable("Hancock", "Female", "<= 2.3")
l5_frequency_tables

cat("===================================================================\n")

analysis_ltable(l5_frequency_tables, 5)
```


```{r L6}
l6_frequency_tables <- generate_ltable("Oklawaha", "Female", "<= 2.3")
l6_frequency_tables

cat("===================================================================\n")

analysis_ltable(l6_frequency_tables, 6)
```


```{r L7}
l7_frequency_tables <- generate_ltable("Trafford", "Female", "<= 2.3")
l7_frequency_tables

cat("===================================================================\n")

analysis_ltable(l7_frequency_tables, 7)
```


```{r L8}
l8_frequency_tables <- generate_ltable("George", "Female", "<= 2.3")
l8_frequency_tables

cat("===================================================================\n")

analysis_ltable(l8_frequency_tables, 8)
```


```{r L9}
l9_frequency_tables <- generate_ltable("Hancock", "Male", "> 2.3")
l9_frequency_tables

cat("===================================================================\n")

analysis_ltable(l9_frequency_tables, 9)
```


```{r L10}
l10_frequency_tables <- generate_ltable("Oklawaha", "Male", "> 2.3")
l10_frequency_tables

cat("===================================================================\n")

analysis_ltable(l10_frequency_tables, 10)
```


```{r L11}
l11_frequency_tables <- generate_ltable("Trafford", "Male", "> 2.3")
l11_frequency_tables

cat("===================================================================\n")

analysis_ltable(l11_frequency_tables, 11)
```


```{r L12}
l12_frequency_tables <- generate_ltable("George", "Male", "> 2.3")
l12_frequency_tables

cat("===================================================================\n")

analysis_ltable(l12_frequency_tables, 12)
```


```{r L13}
l13_frequency_tables <- generate_ltable("Hancock", "Female", "> 2.3")
l13_frequency_tables

cat("===================================================================\n")

analysis_ltable(l13_frequency_tables, 13)
```


```{r L14}
l14_frequency_tables <- generate_ltable("Oklawaha", "Female", "> 2.3")
l14_frequency_tables

cat("===================================================================\n")

analysis_ltable(l14_frequency_tables, 14)
```


```{r L15}
l15_frequency_tables <- generate_ltable("Trafford", "Female", "> 2.3")
l15_frequency_tables

cat("===================================================================\n")

analysis_ltable(l15_frequency_tables, 15)
```


```{r L16}
l16_frequency_tables <- generate_ltable("George", "Female", "> 2.3")
l16_frequency_tables

cat("===================================================================\n")

analysis_ltable(l16_frequency_tables, 16)
```


```{r}
# the degree of freedom in the test above (when Primary Food Type = Bird above) is 
# different from others because there is a "0" in the table

test_table <- as.table(setNames(c(1, 2, 3, 4), 0:3))
test_table2 <- as.table(setNames(c(1, 0, 3, 4), 0:3))

summary(goodfit(test_table, type = "nbinomial"))
summary(goodfit(test_table2, type = "nbinomial"))
```


# Stratified Analysis
```{r measures tables}
Measures <- c("ContCoef", "ContCoefLB", "ContCoefUB", 
              "CramerV", "CramerVLB", "CramerVUB", 
              "Lambda", "LambdaLB", "LambdaUB")

# create tables to store measures 
# name of the table follow pattern: IndependentVariable_BackgroundVariable_table
lake_gender_table <- as.table(array(rep(0, 2*9), dim = c(2, 9), 
                                    dimnames = list("Gender" = Gender, 
                                                    "Measures" = Measures)))

lake_size_table <- as.table(array(rep(0, 2*9), dim = c(2, 9), 
                                  dimnames = list("Size" = Size, 
                                                  "Measures" = Measures)))

gender_lake_table <- as.table(array(rep(0, 4*9), dim = c(4, 9), 
                                    dimnames = list("Lake" = Lake, 
                                                    "Measures" = Measures)))

gender_size_table <- as.table(array(rep(0, 2*9), dim = c(2, 9), 
                                    dimnames = list("Size" = Size, 
                                                    "Measures" = Measures)))

size_lake_table <- as.table(array(rep(0, 4*9), dim = c(4, 9), 
                                  dimnames = list("Lake" = Lake, 
                                                  "Measures" = Measures)))

size_gender_table <- as.table(array(rep(0, 2*9), dim = c(2, 9), 
                                    dimnames = list("Gender" = Gender, 
                                                    "Measures" = Measures)))
```


## Independent Var: Lake; Background Var: Gender
```{r Lake by Gender}

food_choice_table_LbyG <- 
  margin.table(food_choice_table, c("Lake", "Primary_Food_Choice", "Gender"))

# X^2 & G^2 tests
print(assocstats(food_choice_table_LbyG))

# Generalized Cochran-Mantel-Haenszel (CMH) tests
print(CMHtest(food_choice_table_LbyG))

for (gender in dimnames(food_choice_table_LbyG)$Gender){
  LbyG <- food_choice_table_LbyG[,,gender]
  
  # Point Estimate of Contingency Coefficient
  lake_gender_table[gender, "ContCoef"] = round(ContCoef(LbyG), digits = 4)
  
  # Point Estimate of Cramer’s V
  lake_gender_table[gender, "CramerV"] = round(CramerV(LbyG), digits = 4)
  
  # Point Estimate of Goodman Kruskal’s Lambda
  lake_gender_table[gender, "Lambda"] = round(Lambda(LbyG, direction="row"), digits = 4)
  
  d.frm <- Untable(LbyG)
  B <- 1000
  idx <- matrix(sample(nrow(d.frm), size=nrow(d.frm) * B, replace=TRUE), ncol=B, byrow=FALSE)
  
  # code for 95% bootstrap confidence interval for ContCoef
  v_ContCoef <- apply(idx, 2, function(x) ContCoef(d.frm[x,1], d.frm[x,2]))
  v_ContCoef_CI <- quantile(v_ContCoef, probs=c(0.025,0.975), na.rm = TRUE)
  lake_gender_table[gender, "ContCoefLB"] = round(v_ContCoef_CI[[1]], digits = 4)
  lake_gender_table[gender, "ContCoefUB"] = round(v_ContCoef_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for CramerV
  v_CramerV <- apply(idx, 2, function(x) CramerV(d.frm[x,1], d.frm[x,2]))
  v_CramerV_CI <- quantile(v_CramerV, probs=c(0.025,0.975), na.rm = TRUE)
  lake_gender_table[gender, "CramerVLB"] = round(v_CramerV_CI[[1]], digits = 4)
  lake_gender_table[gender, "CramerVUB"] = round(v_CramerV_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for Lambda
  v_Lambda <- apply(idx, 2, function(x) Lambda(d.frm[x,1], d.frm[x,2], direction="row"))
  v_Lambda_CI <- quantile(v_Lambda, probs=c(0.025,0.975), na.rm = TRUE)
  lake_gender_table[gender, "LambdaLB"] = round(v_Lambda_CI[[1]], digits = 4)
  lake_gender_table[gender, "LambdaUB"] = round(v_Lambda_CI[[2]], digits = 4)
}

lake_gender_table
```

The Likelihood Ratio test, Pearson's Chi-squared test, and general test from CMHtest indicate a significant association between lake of capture and primary food type among male alligators and also female alligators, with p-values far below the typical alpha level of 0.05. For male alligotors, Contingency Coefficient and Cramer's V both indicate a moderate association between the lake of capture and primary food type. Goodman Kruskal’s Lambda suggests that knowing the lake of capture can improve the prediction of the primary food type by about 17.33%, which, while modest, is not negligible. For female alligators, the measures of association are very similar to those for males, indicating a moderate association between lake and primary food type among females. Goodman Kruskal’s Lambda for females is higher (23.68%) than for males, suggesting that lake of capture is a slightly better predictor of primary food type for female alligators than for male alligators.




```{r Lake by Gender association, fig.width=8, fig.height=7}
assoc(Primary_Food_Choice ~ Lake + Gender, data = food_choice_table, shade = TRUE)
```


## Independent Var: Lake; Background Var: Size
```{r Lake by Size}

food_choice_table_LbyS <- 
  margin.table(food_choice_table, c("Lake", "Primary_Food_Choice", "Size"))

# X^2 & G^2 tests
print(assocstats(food_choice_table_LbyS))

# Generalized Cochran-Mantel-Haenszel (CMH) tests
print(CMHtest(food_choice_table_LbyS))

for (size in dimnames(food_choice_table_LbyS)$Size){
  LbyS <- food_choice_table_LbyS[,,size]
  
  # Point Estimate of Contingency Coefficient
  lake_size_table[size, "ContCoef"] = round(ContCoef(LbyS), digits = 4)
  
  # Point Estimate of Cramer’s V
  lake_size_table[size, "CramerV"] = round(CramerV(LbyS), digits = 4)
  
  # Point Estimate of Goodman Kruskal’s Lambda
  lake_size_table[size, "Lambda"] = round(Lambda(LbyS, direction="row"), digits = 4)
  
  d.frm <- Untable(LbyS)
  B <- 1000
  idx <- matrix(sample(nrow(d.frm), size=nrow(d.frm) * B, replace=TRUE), ncol=B, byrow=FALSE)
  
  # code for 95% bootstrap confidence interval for ContCoef
  v_ContCoef <- apply(idx, 2, function(x) ContCoef(d.frm[x,1], d.frm[x,2]))
  v_ContCoef_CI <- quantile(v_ContCoef, probs=c(0.025,0.975), na.rm = TRUE)
  lake_size_table[size, "ContCoefLB"] = round(v_ContCoef_CI[[1]], digits = 4)
  lake_size_table[size, "ContCoefUB"] = round(v_ContCoef_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for CramerV
  v_CramerV <- apply(idx, 2, function(x) CramerV(d.frm[x,1], d.frm[x,2]))
  v_CramerV_CI <- quantile(v_CramerV, probs=c(0.025,0.975), na.rm = TRUE)
  lake_size_table[size, "CramerVLB"] = round(v_CramerV_CI[[1]], digits = 4)
  lake_size_table[size, "CramerVUB"] = round(v_CramerV_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for Lambda
  v_Lambda <- apply(idx, 2, function(x) Lambda(d.frm[x,1], d.frm[x,2], direction="row"))
  v_Lambda_CI <- quantile(v_Lambda, probs=c(0.025,0.975), na.rm = TRUE)
  lake_size_table[size, "LambdaLB"] = round(v_Lambda_CI[[1]], digits = 4)
  lake_size_table[size, "LambdaUB"] = round(v_Lambda_CI[[2]], digits = 4)
}

lake_size_table
```


```{r Lake by Size association, fig.width=8, fig.height=7}
assoc(Primary_Food_Choice ~ Lake + Size, data = food_choice_table, shade = TRUE)
```


## Independent Var: Gender; Background Var: Lake
```{r Gender by Lake}

food_choice_table_GbyL <- 
  margin.table(food_choice_table, c("Gender", "Primary_Food_Choice", "Lake"))

# X^2 & G^2 tests
print(assocstats(food_choice_table_GbyL))

# Generalized Cochran-Mantel-Haenszel (CMH) tests
print(CMHtest(food_choice_table_GbyL))

for (lake in dimnames(food_choice_table_GbyL)$Lake){
  GbyL <- food_choice_table_GbyL[,,lake]
  
  # Point Estimate of Contingency Coefficient
  gender_lake_table[lake, "ContCoef"] = round(ContCoef(GbyL), digits = 4)
  
  # Point Estimate of Cramer’s V
  gender_lake_table[lake, "CramerV"] = round(CramerV(GbyL), digits = 4)
  
  # Point Estimate of Goodman Kruskal’s Lambda
  gender_lake_table[lake, "Lambda"] = round(Lambda(GbyL, direction="row"), digits = 4)
  
  d.frm <- Untable(GbyL)
  B <- 1000
  idx <- matrix(sample(nrow(d.frm), size=nrow(d.frm) * B, replace=TRUE), ncol=B, byrow=FALSE)
  
  # code for 95% bootstrap confidence interval for ContCoef
  v_ContCoef <- apply(idx, 2, function(x) ContCoef(d.frm[x,1], d.frm[x,2]))
  v_ContCoef_CI <- quantile(v_ContCoef, probs=c(0.025,0.975), na.rm = TRUE)
  gender_lake_table[lake, "ContCoefLB"] = round(v_ContCoef_CI[[1]], digits = 4)
  gender_lake_table[lake, "ContCoefUB"] = round(v_ContCoef_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for CramerV
  v_CramerV <- apply(idx, 2, function(x) CramerV(d.frm[x,1], d.frm[x,2]))
  v_CramerV_CI <- quantile(v_CramerV, probs=c(0.025,0.975), na.rm = TRUE)
  gender_lake_table[lake, "CramerVLB"] = round(v_CramerV_CI[[1]], digits = 4)
  gender_lake_table[lake, "CramerVUB"] = round(v_CramerV_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for Lambda
  v_Lambda <- apply(idx, 2, function(x) Lambda(d.frm[x,1], d.frm[x,2], direction="row"))
  v_Lambda_CI <- quantile(v_Lambda, probs=c(0.025,0.975), na.rm = TRUE)
  gender_lake_table[lake, "LambdaLB"] = round(v_Lambda_CI[[1]], digits = 4)
  gender_lake_table[lake, "LambdaUB"] = round(v_Lambda_CI[[2]], digits = 4)
}

gender_lake_table
```


```{r Gender by Lake association, fig.width=8, fig.height=7}
food_choice_table_lake_abbrev <- food_choice_table
dimnames(food_choice_table_lake_abbrev)["Lake"] <- list(c("H", "O", "T", "G"))
assoc(Primary_Food_Choice ~ Gender + Lake, data = food_choice_table_lake_abbrev, shade = TRUE)
```


## Independent Var: Gender; Background Var: Size
```{r Gender by Size}

food_choice_table_GbyS <- 
  margin.table(food_choice_table, c("Gender", "Primary_Food_Choice", "Size"))

# X^2 & G^2 tests
print(assocstats(food_choice_table_GbyS))

# Generalized Cochran-Mantel-Haenszel (CMH) tests
print(CMHtest(food_choice_table_GbyS))

for (size in dimnames(food_choice_table_GbyS)$Size){
  GbyS <- food_choice_table_GbyS[,,size]
  
  # Point Estimate of Contingency Coefficient
  gender_size_table[size, "ContCoef"] = round(ContCoef(GbyS), digits = 4)
  
  # Point Estimate of Cramer’s V
  gender_size_table[size, "CramerV"] = round(CramerV(GbyS), digits = 4)
  
  # Point Estimate of Goodman Kruskal’s Lambda
  gender_size_table[size, "Lambda"] = round(Lambda(GbyS, direction="row"), digits = 4)
  
  d.frm <- Untable(GbyS)
  B <- 1000
  idx <- matrix(sample(nrow(d.frm), size=nrow(d.frm) * B, replace=TRUE), ncol=B, byrow=FALSE)
  
  # code for 95% bootstrap confidence interval for ContCoef
  v_ContCoef <- apply(idx, 2, function(x) ContCoef(d.frm[x,1], d.frm[x,2]))
  v_ContCoef_CI <- quantile(v_ContCoef, probs=c(0.025,0.975), na.rm = TRUE)
  gender_size_table[size, "ContCoefLB"] = round(v_ContCoef_CI[[1]], digits = 4)
  gender_size_table[size, "ContCoefUB"] = round(v_ContCoef_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for CramerV
  v_CramerV <- apply(idx, 2, function(x) CramerV(d.frm[x,1], d.frm[x,2]))
  v_CramerV_CI <- quantile(v_CramerV, probs=c(0.025,0.975), na.rm = TRUE)
  gender_size_table[size, "CramerVLB"] = round(v_CramerV_CI[[1]], digits = 4)
  gender_size_table[size, "CramerVUB"] = round(v_CramerV_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for Lambda
  v_Lambda <- apply(idx, 2, function(x) Lambda(d.frm[x,1], d.frm[x,2], direction="row"))
  v_Lambda_CI <- quantile(v_Lambda, probs=c(0.025,0.975), na.rm = TRUE)
  gender_size_table[size, "LambdaLB"] = round(v_Lambda_CI[[1]], digits = 4)
  gender_size_table[size, "LambdaUB"] = round(v_Lambda_CI[[2]], digits = 4)
}

gender_size_table
```


```{r Gender by Size association, fig.width=8, fig.height=7}
assoc(Primary_Food_Choice ~ Gender + Size, data = food_choice_table, shade = TRUE)
```


## Independent Var: Size; Background Var: Lake
```{r Size by Lake}

food_choice_table_SbyL <- 
  margin.table(food_choice_table, c("Size", "Primary_Food_Choice", "Lake"))

# X^2 & G^2 tests
print(assocstats(food_choice_table_SbyL))

# Generalized Cochran-Mantel-Haenszel (CMH) tests
print(CMHtest(food_choice_table_SbyL))

for (lake in dimnames(food_choice_table_SbyL)$Lake){
  SbyL <- food_choice_table_SbyL[,,lake]
  
  # Point Estimate of Contingency Coefficient
  size_lake_table[lake, "ContCoef"] = round(ContCoef(SbyL), digits = 4)
  
  # Point Estimate of Cramer’s V
  size_lake_table[lake, "CramerV"] = round(CramerV(SbyL), digits = 4)
  
  # Point Estimate of Goodman Kruskal’s Lambda
  size_lake_table[lake, "Lambda"] = round(Lambda(SbyL, direction="row"), digits = 4)
  
  d.frm <- Untable(SbyL)
  B <- 1000
  idx <- matrix(sample(nrow(d.frm), size=nrow(d.frm) * B, replace=TRUE), ncol=B, byrow=FALSE)
  
  # code for 95% bootstrap confidence interval for ContCoef
  v_ContCoef <- apply(idx, 2, function(x) ContCoef(d.frm[x,1], d.frm[x,2]))
  v_ContCoef_CI <- quantile(v_ContCoef, probs=c(0.025,0.975), na.rm = TRUE)
  size_lake_table[lake, "ContCoefLB"] = round(v_ContCoef_CI[[1]], digits = 4)
  size_lake_table[lake, "ContCoefUB"] = round(v_ContCoef_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for CramerV
  v_CramerV <- apply(idx, 2, function(x) CramerV(d.frm[x,1], d.frm[x,2]))
  v_CramerV_CI <- quantile(v_CramerV, probs=c(0.025,0.975), na.rm = TRUE)
  size_lake_table[lake, "CramerVLB"] = round(v_CramerV_CI[[1]], digits = 4)
  size_lake_table[lake, "CramerVUB"] = round(v_CramerV_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for Lambda
  v_Lambda <- apply(idx, 2, function(x) Lambda(d.frm[x,1], d.frm[x,2], direction="row"))
  v_Lambda_CI <- quantile(v_Lambda, probs=c(0.025,0.975), na.rm = TRUE)
  size_lake_table[lake, "LambdaLB"] = round(v_Lambda_CI[[1]], digits = 4)
  size_lake_table[lake, "LambdaUB"] = round(v_Lambda_CI[[2]], digits = 4)
}

size_lake_table
```


```{r Size by Lake association, fig.width=8, fig.height=7}
food_choice_table_lake_abbrev <- food_choice_table
dimnames(food_choice_table_lake_abbrev)["Lake"] <- list(c("H", "O", "T", "G"))
assoc(Primary_Food_Choice ~ Size + Lake, data = food_choice_table_lake_abbrev, shade = TRUE)
```


## Independent Var: Size; Background Var: Gender
```{r Size by Gender}

food_choice_table_SbyG <- 
  margin.table(food_choice_table, c("Size", "Primary_Food_Choice", "Gender"))

# X^2 & G^2 tests
print(assocstats(food_choice_table_SbyG))

# Generalized Cochran-Mantel-Haenszel (CMH) tests
print(CMHtest(food_choice_table_SbyG))

for (gender in dimnames(food_choice_table_SbyG)$Gender){
  SbyG <- food_choice_table_SbyG[,,gender]
  
  # Point Estimate of Contingency Coefficient
  size_gender_table[gender, "ContCoef"] = round(ContCoef(SbyG), digits = 4)
  
  # Point Estimate of Cramer’s V
  size_gender_table[gender, "CramerV"] = round(CramerV(SbyG), digits = 4)
  
  # Point Estimate of Goodman Kruskal’s Lambda
  size_gender_table[gender, "Lambda"] = round(Lambda(SbyG, direction="row"), digits = 4)
  
  d.frm <- Untable(SbyG)
  B <- 1000
  idx <- matrix(sample(nrow(d.frm), size=nrow(d.frm) * B, replace=TRUE), ncol=B, byrow=FALSE)
  
  # code for 95% bootstrap confidence interval for ContCoef
  v_ContCoef <- apply(idx, 2, function(x) ContCoef(d.frm[x,1], d.frm[x,2]))
  v_ContCoef_CI <- quantile(v_ContCoef, probs=c(0.025,0.975), na.rm = TRUE)
  size_gender_table[gender, "ContCoefLB"] = round(v_ContCoef_CI[[1]], digits = 4)
  size_gender_table[gender, "ContCoefUB"] = round(v_ContCoef_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for CramerV
  v_CramerV <- apply(idx, 2, function(x) CramerV(d.frm[x,1], d.frm[x,2]))
  v_CramerV_CI <- quantile(v_CramerV, probs=c(0.025,0.975), na.rm = TRUE)
  size_gender_table[gender, "CramerVLB"] = round(v_CramerV_CI[[1]], digits = 4)
  size_gender_table[gender, "CramerVUB"] = round(v_CramerV_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for Lambda
  v_Lambda <- apply(idx, 2, function(x) Lambda(d.frm[x,1], d.frm[x,2], direction="row"))
  v_Lambda_CI <- quantile(v_Lambda, probs=c(0.025,0.975), na.rm = TRUE)
  size_gender_table[gender, "LambdaLB"] = round(v_Lambda_CI[[1]], digits = 4)
  size_gender_table[gender, "LambdaUB"] = round(v_Lambda_CI[[2]], digits = 4)
}

size_gender_table
```


```{r Size by Gender association, fig.width=8, fig.height=7}
assoc(Primary_Food_Choice ~ Size + Gender, data = food_choice_table, shade = TRUE)
```


## Association Plots
```{r association plot, fig.width=10, fig.height=16}
assocL <- assoc(Primary_Food_Choice ~ Lake, data = food_choice_table, shade = TRUE)

assocG <- assoc(Primary_Food_Choice ~ Gender, data = food_choice_table, shade = TRUE)

assocS <- assoc(Primary_Food_Choice ~ Size, data = food_choice_table, shade = TRUE)

assocLG <- assoc(Primary_Food_Choice ~ Lake + Gender, data = food_choice_table, shade = TRUE)

assocLS <- assoc(Primary_Food_Choice ~ Lake + Size, data = food_choice_table, shade = TRUE)

assocGS <- assoc(Primary_Food_Choice ~ Gender + Size, data = food_choice_table, shade = TRUE)

assocLGS <- assoc(Primary_Food_Choice ~ Lake + Gender + Size, data = food_choice_table, shade = TRUE)

assoc(~ Lake + Primary_Food_Choice + Gender + Size, data = food_choice_table, shade = TRUE)
```


# Doubledecker Plot
```{r, fig.width=16, fig.height=8}
food_choice_tableDD <- food_choice_table
dimnames(food_choice_tableDD)["Gender"] <- list(c("M", "F"))
doubledecker(Primary_Food_Choice ~ Lake + Size + Gender, data = food_choice_tableDD)
```




























# Failed Cases
```{r Size by Lake_Gender}

# Combine Lake and Gender
Lake_Gender <- outer(Lake, Gender, paste, sep = "_")

# Transpose and reshape the original array for Lake_Gender combination
food_choice_array_LG <- aperm(food_choice_array, c(3, 4, 1, 2))
dim(food_choice_array_LG) <- c(
  length(Size), 
  length(Primary_Food_Choice), 
  length(Lake) * length(Gender)
)

# Assign dimension names
dimnames(food_choice_array_LG) <- list(
  Size = Size,
  Primary_Food_Choice = Primary_Food_Choice,
  Lake_Gender = as.vector(Lake_Gender)
)

# Convert to table
food_choice_table_LG <- as.table(food_choice_array_LG)
# food_choice_table_LG
# dimnames(food_choice_table_LG)

Size_LakeGender_table <- as.table(array(rep(0, 8*9), dim = c(8, 9), 
                                    dimnames = list("Lake_Gender" = Lake_Gender, 
                                                    "Measures" = Measures)))


# X^2 & G^2 tests
print(assocstats(food_choice_table_LG))

# Generalized Cochran-Mantel-Haenszel (CMH) tests
# print(CMHtest(food_choice_table_LG))

for (lake_gender in dimnames(food_choice_table_LG)$Lake_Gender){
  SbyLG <- food_choice_table_LG[,,lake_gender]
  
  # Point Estimate of Contingency Coefficient
  Size_LakeGender_table[lake_gender, "ContCoef"] = round(ContCoef(SbyLG), digits = 4)
  
  # Point Estimate of Cramer’s V
  Size_LakeGender_table[lake_gender, "CramerV"] = round(CramerV(SbyLG), digits = 4)
  
  # Point Estimate of Goodman Kruskal’s Lambda
  Size_LakeGender_table[lake_gender, "Lambda"] = round(Lambda(SbyLG, direction="row"), digits = 4)
  
  d.frm <- Untable(SbyLG)
  B <- 1000
  idx <- matrix(sample(nrow(d.frm), size=nrow(d.frm) * B, replace=TRUE), ncol=B, byrow=FALSE)
  
  # code for 95% bootstrap confidence interval for ContCoef
  v_ContCoef <- apply(idx, 2, function(x) ContCoef(d.frm[x,1], d.frm[x,2]))
  v_ContCoef_CI <- quantile(v_ContCoef, probs=c(0.025,0.975), na.rm = TRUE)
  Size_LakeGender_table[lake_gender, "ContCoefLB"] = round(v_ContCoef_CI[[1]], digits = 4)
  Size_LakeGender_table[lake_gender, "ContCoefUB"] = round(v_ContCoef_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for CramerV
  v_CramerV <- apply(idx, 2, function(x) CramerV(d.frm[x,1], d.frm[x,2]))
  v_CramerV_CI <- quantile(v_CramerV, probs=c(0.025,0.975), na.rm = TRUE)
  Size_LakeGender_table[lake_gender, "CramerVLB"] = round(v_CramerV_CI[[1]], digits = 4)
  Size_LakeGender_table[lake_gender, "CramerVUB"] = round(v_CramerV_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for Lambda
  v_Lambda <- apply(idx, 2, function(x) Lambda(d.frm[x,1], d.frm[x,2], direction="row"))
  v_Lambda_CI <- quantile(v_Lambda, probs=c(0.025,0.975), na.rm = TRUE)
  Size_LakeGender_table[lake_gender, "LambdaLB"] = round(v_Lambda_CI[[1]], digits = 4)
  Size_LakeGender_table[lake_gender, "LambdaUB"] = round(v_Lambda_CI[[2]], digits = 4)
}

Size_LakeGender_table

```



```{r Gender by Lake_Size}

# Combine Lake and Size
Lake_Size <- outer(Lake, Size, paste, sep = "_")

# Transpose and reshape the original array for Lake_Size combination
food_choice_array_LS <- aperm(food_choice_array, c(2, 4, 1, 3))
dim(food_choice_array_LS) <- c(
  length(Gender), 
  length(Primary_Food_Choice), 
  length(Lake) * length(Size)
)

# Assign dimension names
dimnames(food_choice_array_LS) <- list(
  Gender = Gender,
  Primary_Food_Choice = Primary_Food_Choice,
  Lake_Size = as.vector(Lake_Size)
)

# Convert to table
food_choice_table_LS <- as.table(food_choice_array_LS)
# food_choice_table_LS
# dimnames(food_choice_table_LS)

Gender_LakeSize_table <- as.table(array(rep(0, 8*9), dim = c(8, 9),
                                    dimnames = list("Lake_Size" = Lake_Size,
                                                    "Measures" = Measures)))


# X^2 & G^2 tests
print(assocstats(food_choice_table_LS))

# Generalized Cochran-Mantel-Haenszel (CMH) tests
# print(CMHtest(food_choice_table_LS))

for (lake_size in dimnames(food_choice_table_LS)$Lake_Size){
  GbyLS <- food_choice_table_LS[,,lake_size]
  
  # Point Estimate of Contingency Coefficient
  Gender_LakeSize_table[lake_size, "ContCoef"] = round(ContCoef(GbyLS), digits = 4)
  
  # Point Estimate of Cramer’s V
  Gender_LakeSize_table[lake_size, "CramerV"] = round(CramerV(GbyLS), digits = 4)
  
  # Point Estimate of Goodman Kruskal’s Lambda
  Gender_LakeSize_table[lake_size, "Lambda"] = round(Lambda(GbyLS, direction="row"), digits = 4)
  
  d.frm <- Untable(GbyLS)
  B <- 1000
  idx <- matrix(sample(nrow(d.frm), size=nrow(d.frm) * B, replace=TRUE), ncol=B, byrow=FALSE)
  
  # code for 95% bootstrap confidence interval for ContCoef
  v_ContCoef <- apply(idx, 2, function(x) ContCoef(d.frm[x,1], d.frm[x,2]))
  v_ContCoef_CI <- quantile(v_ContCoef, probs=c(0.025,0.975), na.rm = TRUE)
  Gender_LakeSize_table[lake_size, "ContCoefLB"] = round(v_ContCoef_CI[[1]], digits = 4)
  Gender_LakeSize_table[lake_size, "ContCoefUB"] = round(v_ContCoef_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for CramerV
  v_CramerV <- apply(idx, 2, function(x) CramerV(d.frm[x,1], d.frm[x,2]))
  v_CramerV_CI <- quantile(v_CramerV, probs=c(0.025,0.975), na.rm = TRUE)
  Gender_LakeSize_table[lake_size, "CramerVLB"] = round(v_CramerV_CI[[1]], digits = 4)
  Gender_LakeSize_table[lake_size, "CramerVUB"] = round(v_CramerV_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for Lambda
  v_Lambda <- apply(idx, 2, function(x) Lambda(d.frm[x,1], d.frm[x,2], direction="row"))
  v_Lambda_CI <- quantile(v_Lambda, probs=c(0.025,0.975), na.rm = TRUE)
  Gender_LakeSize_table[lake_size, "LambdaLB"] = round(v_Lambda_CI[[1]], digits = 4)
  Gender_LakeSize_table[lake_size, "LambdaUB"] = round(v_Lambda_CI[[2]], digits = 4)
}

Gender_LakeSize_table

```



```{r Lake by Gender_Size}

# Combine Gender and Size
Gender_Size <- outer(Gender, Size, paste, sep = "_")

# Transpose and reshape the original array for Gender_Size combination
food_choice_array_GS <- aperm(food_choice_array, c(1, 4, 2, 3))
dim(food_choice_array_GS) <- c(
  length(Lake), 
  length(Primary_Food_Choice), 
  length(Gender) * length(Size)
)

# Assign dimension names
dimnames(food_choice_array_GS) <- list(
  Lake = Lake,
  Primary_Food_Choice = Primary_Food_Choice,
  Gender_Size = as.vector(Gender_Size)
)

# Convert to table
food_choice_table_GS <- as.table(food_choice_array_GS)
# food_choice_table_GS
# dimnames(food_choice_table_GS)

Lake_GenderSize_table <- as.table(array(rep(0, 4*9), dim = c(4, 9),
                                    dimnames = list("Gender_Size" = Gender_Size,
                                                    "Measures" = Measures)))


# X^2 & G^2 tests
print(assocstats(food_choice_table_GS))

# Generalized Cochran-Mantel-Haenszel (CMH) tests
# print(CMHtest(food_choice_table_GS))

for (gender_size in dimnames(food_choice_table_GS)$Gender_Size){
  LbyGS <- food_choice_table_GS[,,gender_size]
  
  # Point Estimate of Contingency Coefficient
  Lake_GenderSize_table[gender_size, "ContCoef"] = round(ContCoef(LbyGS), digits = 4)
  
  # Point Estimate of Cramer’s V
  Lake_GenderSize_table[gender_size, "CramerV"] = round(CramerV(LbyGS), digits = 4)
  
  # Point Estimate of Goodman Kruskal’s Lambda
  Lake_GenderSize_table[gender_size, "Lambda"] = round(Lambda(LbyGS, direction="row"), digits = 4)
  
  d.frm <- Untable(LbyGS)
  B <- 1000
  idx <- matrix(sample(nrow(d.frm), size=nrow(d.frm) * B, replace=TRUE), ncol=B, byrow=FALSE)
  
  # code for 95% bootstrap confidence interval for ContCoef
  v_ContCoef <- apply(idx, 2, function(x) ContCoef(d.frm[x,1], d.frm[x,2]))
  v_ContCoef_CI <- quantile(v_ContCoef, probs=c(0.025,0.975), na.rm = TRUE)
  Lake_GenderSize_table[gender_size, "ContCoefLB"] = round(v_ContCoef_CI[[1]], digits = 4)
  Lake_GenderSize_table[gender_size, "ContCoefUB"] = round(v_ContCoef_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for CramerV
  v_CramerV <- apply(idx, 2, function(x) CramerV(d.frm[x,1], d.frm[x,2]))
  v_CramerV_CI <- quantile(v_CramerV, probs=c(0.025,0.975), na.rm = TRUE)
  Lake_GenderSize_table[gender_size, "CramerVLB"] = round(v_CramerV_CI[[1]], digits = 4)
  Lake_GenderSize_table[gender_size, "CramerVUB"] = round(v_CramerV_CI[[2]], digits = 4)
  
  # code for 95% bootstrap confidence interval for Lambda
  v_Lambda <- apply(idx, 2, function(x) Lambda(d.frm[x,1], d.frm[x,2], direction="row"))
  v_Lambda_CI <- quantile(v_Lambda, probs=c(0.025,0.975), na.rm = TRUE)
  Lake_GenderSize_table[gender_size, "LambdaLB"] = round(v_Lambda_CI[[1]], digits = 4)
  Lake_GenderSize_table[gender_size, "LambdaUB"] = round(v_Lambda_CI[[2]], digits = 4)
}

Lake_GenderSize_table

```







